Untitled

### 2022-12-21 draft1_Rcode.Rmd
Updating final R code to include Q2 code and part b and c from Q3

```{r}
## Question 2
n <- 100000
X <- rexp(n,rate=0.5)+30 ## generate samples from shifted Exp(10/7)
v.weights <- dnorm(X,mean=30,sd=0.7)/dexp(X-30,rate=0.5) ## vector of weights
p.t <- mean(v.weights) ## calculate the IS estimator

q2.is <- function(n,x0,time,a,lambda){

}
```

>>>>

```{r}
## Question 2

# Trial function calculates probability of X_1 >a_1 given initial value (need to now work out how to do the rest)
q2.is.trial <- function(n,x0,time,a,lambda){
  #Turning values into proportions
  a <- a/100
  x0 <- x0/100
  #transforming a to logit scale and turning a into a proportion (dividing by 100)
  b <- log((a)/(1-a))
  #generating samples from shifted exponential
  X <- rexp(n,rate = lambda) + (b-x0)/sqrt(time[2]-time[1])
  #Calculating weights
  v.weights <- dnorm(X)/dexp(X-((b-x0)/sqrt(time[2]-time[1])),rate=lambda)
  #Finding I.S estimator
  p.t <- mean(v.weights)
  return(p.t)
}

q2.is.trial(n=10000,x0=30,time=c(0,0.7),a=50,lambda=0.5)

q2.is.trial.chain <- function(n,x0,time,a,lambda){
  #Turning values into proportions
  a <- a/100
  x0 <- x0/100
  #transforming a to logit scale (calling this vector b)
  b <- log((a)/(1-a))

  p.b.iter <- NULL

  if (length(time)>2){  #Case when Markov chain has more than 1 transition thus have to simulate a random walk to calculate the proposal distribution

    for (k in 1:n){

      #Simulating random walk
      r.w <- NULL
      r.w[1] <- x0
      for (i in 2:(length(time)-1)){
        r.w[i] <- rnorm(1,mean=r.w[i-1],sd=sqrt(time[i]-time[i-1]))
      }

      #generating samples from shifted exponential
      X <- NULL
      for (i in 2:length(time)){
        X[i-1] <- rexp(1,rate = lambda) + (b[i-1]-r.w[i-1])/sqrt(time[i]-time[i-1])
      }
      #Calculating weights for each observation in random walk
      v.weights <- NULL
      for (i in 2:length(time)){
        v.weights[i-1] <- dnorm(X[i-1])/dexp(X[i-1]-((b[i-1]-r.w[i-1])/sqrt(time[i]-time[i-1])),rate=lambda)
      }
      #Finding probability that X is greater than a (satisfies equation 6) for current iteration 
      p.b.iter[k] <- prod(v.weights)
      }
      #Finding I.S estimator by taking mean 
      I.S.est <- mean(p.b.iter)

  } else { # Dealing with case that markov chain only has 1 jump so don't need to calculate random walk
           # Simplifies to case that X_{1}>a given that X_{0}=x0

    #generating samples from proposal shifted exponential
    X <- rexp(n,rate = lambda) + (b-x0)/sqrt(time[2]-time[1])
    #Calculating weights
    v.weights <- dnorm(X)/dexp(X-((b-x0)/sqrt(time[2]-time[1])),rate=lambda)
    #Finding I.S estimator
    I.S.est  <- mean(v.weights)
  }

  return(I.S.est)
}
set.seed(2022)
q2.is.trial.chain(n=10000,x0=30,time=c(0,0.7,1.2,2.5),a=c(50,70,90),lambda=0.5)
set.seed(2022)
q2.is.trial.chain(n=10000,x0=30,time=c(0,0.7),a=c(50),lambda=0.5)

set.seed(2022)
n=10000
x0=5
time=c(0,0.7,1.2,2.5)
a=c(50,70,90)
lambda=0.5

  a <- a/100
  x0 <- x0/100
  #transforming a to logit scale and turning a into a proportion (dividing by 100)
  b <- log((a)/(1-a))
  p.t <- NULL
  for (k in 1:n){

  #Simulating random walk
  r.w <- NULL
  r.w[1] <- x0
  for (i in 2:(length(time)-1)){
    r.w[i] <- rnorm(1,mean=r.w[i-1],sd=sqrt(time[i]-time[i-1]))
  }

  #generating samples from shifted exponential
  X <- NULL
  for (i in 2:length(time)){
  X[i-1] <- rexp(1,rate = lambda) + (b[i-1]-r.w[i-1])/sqrt(time[i]-time[i-1])
  }
  #Calculating weights for each observation in random walk
  v.weights <- NULL
  for (i in 2:length(time)){
  v.weights[i-1] <- dnorm(X[i-1])/dexp(X[i-1]-((b[i-1]-r.w[i-1])/sqrt(time[i]-time[i-1])),rate=lambda)
  }
  #Finding I.S estimator (product means to find overall)
  p.t[k] <- prod(mean(v.weights))
  }

n <- 100
x0 <- 30
time <- c(0,0.7)
a <- 50
lambda <- 0.5

  b <- log((a/100)/(1-(a/100)))
  #generating samples from shifted exponential
  X <- rexp(n,rate = lambda) + (b-x0)/sqrt(time[2]-time[1])
  v.weights <- dnorm(X)/dexp(X-((b-x0)/sqrt(time[2]-time[1])),rate=lambda)
  p.t <- mean(v.weights)
  return(p.t)
```
